The imprint of global magma oceans on exoplanet demographics

### 2023-03-09 src/tex/ms.tex
describe overall method earlier

\section{Baseline Model}\label{sec:met_baseline}
%The purpose of this paper is to demonstrate the feasibility of measuring the demographic feature introduced above with near-future exoplanet missions.
\todo[inline]{Start methods description with brief statement of goal and overview of the steps. The first sentence is similar, but would be good to separate goal from method here.
}
We determined -- for different survey configurations -- the confidence level with which the runaway greenhouse threshold can be detected in simulated exoplanet observations.
For this purpose we expanded the \bioverse\ framework~\citep{Bixel2020,Bixel2021}\footnote{\bioverse\ is actively maintained and documented open source software written in Python. Its latest version and documentation can be found at \url{https://github.com/danielapai/bioverse}.} to generate synthetic stellar and planetary samples into which we injected the runaway greenhouse signal, simulated observations of this sample, and computed Bayesian evidences in favor of the runaway greenhouse hypothesis (see diagram in Fig.~\ref{fig:flowchart}).
\todo[inline]{\bioverse\ will need a short introduction before the modifications are mentioned. Probably should get a paragraph in the intro}
\begin{figure*}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/flowchart.pdf}
        \caption{Workflow of our hypothesis testing with \bioverse. In the first block, we generate a sample of stars and populate them with planets based on \kepler\ demographics.
            A fraction of them are then assigned a runaway greenhouse climate based on the model described in Sect.~\ref{sec:met_rghmodel}.
        The second block simulates an exoplanet survey, whereby selection effects and detection biases are introduced. Finally, the third block deals with testing the runaway greenhouse hypothesis based on data from the survey simulation.
        By iterating through these steps, we compute the statistical power of testing the hypothesis given the assumed survey design.}
        \label{fig:flowchart}
    \end{centering}
\end{figure*}
In the following, we review the source of the stellar sample, the modeled luminosity evolution, the generation of a synthetic planet sample, as well as the orbital parameters of the planets.

@@

Central to our procedure is to explore the detectability of population-level trends with future exoplanet surveys, and here the trend in question is the proposed radius discontinuity.
To enable a quantitative assessment of this detectability, we injected the signal into the simulated planet sample before observing it with simulated surveys.
While we search for the runaway greenhouse pattern in demographic quantities such as average planet radii, the injected changes happen on the planetary level: We changed each planet's radius based on its individual set of properties and the associated predictions from steam atmosphere and water incorporation models.
Relevant properties are a planet's mass $M$, its net instellation $S$, and its bulk water inventory expressed as a water mass fraction $x_{H_2O}$.
We consider the following cases (see Fig.~\ref{fig:radiusevolution}):
%\begin{itemize}
%\item[A.] No radius change.
%This case serves as our null hypothesis, where planetary radii are not affected by the amount of irradiation the planets receive.
%A solid planetary surface is assumed and no steam atmosphere is present.
%\item[B.] Radius inflation due to a steam atmosphere.
%Atmospheric scale heights are significantly increased.
%\item[C.] As B, but including radius decrease of inflated planets due to water partitioning in the molten mantle.
%\end{itemize}

@@

\subsubsection{Spectral type of target stars}\label{sec:results_FGK_M}
Since the receiving radiation is sensitive on the spectral type of the host star, the share of planets on either side of $S_\mathrm{thresh}$ is different for FGK and M~dwarfs.
We tested the detectability of the runaway greenhouse signal when only FGK or only M~dwarfs are considered (see Fig.~\ref{fig:plato_fgrid}).
%($d_\mathrm{max} = \SI{75}{\parsec}, M_\mathrm{G} < 11$)
The samples are volume and magnitude-limited to reflect the target counts of \plato's provisional Long-duration Observation Phase fields~\citep[$15996$ FGK stars in the P1 and P2 samples, $33948$ M~stars in the P4 sample, ][]{Nascimbeni2022}.
The resulting M~dwarf planet sample is significantly larger with $\var{N_M}\pm \var{N_M_err}$ planets compared to $\var{N_FGK}\pm \var{N_FGK_err}$ planets in the FGK sample.
No significant detection is possible in the pure FGK sample, independent of the assumed geophysical parameters.
In the M~dwarf sample, the evidence threshold is reached around $f_\mathrm{rgh} \sim 0.2$, similar to the case above where all spectral types are considered.

>>>>

\section{Baseline Model}\label{sec:met_baseline}
%The purpose of this paper is to demonstrate the feasibility of measuring the demographic feature introduced above with near-future exoplanet missions.
\todo[inline]{Start methods description with brief statement of goal and overview of the steps. The first sentence is similar, but would be good to separate goal from method here.
}
We determined -- for different survey configurations -- the confidence level with which the runaway greenhouse threshold can be detected in simulated exoplanet observations.
For this purpose we expanded the \bioverse\ framework~\citep{Bixel2020,Bixel2021}\footnote{\bioverse\ is actively maintained and documented open source software written in Python. Its latest version and documentation can be found at \url{https://github.com/danielapai/bioverse}.} to generate synthetic stellar and planetary samples into which we injected the runaway greenhouse signal, simulated observations of this sample, and computed Bayesian evidences in favor of the runaway greenhouse hypothesis (see diagram in Fig.~\ref{fig:flowchart}).
\todo[inline]{\bioverse\ will need a short introduction before the modifications are mentioned. Probably should get a paragraph in the intro}
\begin{figure*}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/flowchart.pdf}
        \caption{Workflow of our hypothesis testing with \bioverse. First, we generate a sample of stars and populate them with planets based on \kepler\ demographics.
            A fraction of them are then assigned a runaway greenhouse climate based on the model described in Sect.~\ref{sec:met_rghmodel}.
        We then simulate an exoplanet survey, whereby selection effects and detection biases are introduced. Finally, we test the runaway greenhouse hypothesis based on data from the survey simulation.
        By iterating through these steps, we compute the statistical power of testing the hypothesis for different survey designs.}
        \label{fig:flowchart}
    \end{centering}
\end{figure*}
Central to our procedure is to explore the detectability of population-level trends with future exoplanet surveys, and here the trend in question is the proposed radius discontinuity.
To enable a quantitative assessment of this detectability, we injected the signal into a simulated planet sample before observing it with simulated surveys.
In the following, we review the source of the stellar sample, the modeled luminosity evolution, the generation of a synthetic planet sample, as well as the orbital parameters of the planets.

@@

While we search for the runaway greenhouse pattern in demographic quantities such as average planet radii, the injected changes happen on the planetary level: We changed each planet's radius based on its individual set of properties and the associated predictions from steam atmosphere and water incorporation models.
Relevant properties are a planet's mass $M$, its net instellation $S$, and its bulk water inventory expressed as a water mass fraction $x_{H_2O}$.
We consider the following cases (see Fig.~\ref{fig:radiusevolution}):

@@

\subsubsection{Dependence on host star spectral type}\label{sec:results_FGK_M}
Since the receiving radiation is sensitive on the spectral type of the host star, the share of planets on either side of $S_\mathrm{thresh}$ is different for FGK and M~dwarfs.
We tested the detectability of the runaway greenhouse signal when only FGK or only M~dwarfs are considered (see Fig.~\ref{fig:plato_fgrid}).
%($d_\mathrm{max} = \SI{75}{\parsec}, M_\mathrm{G} < 11$)
The samples are volume and magnitude-limited to reflect the target counts of \plato's provisional Long-duration Observation Phase fields~\citep[$15996$ FGK stars in the P1 and P2 samples, $33948$ M~stars in the P4 sample, ][]{Nascimbeni2022}.
The resulting M~dwarf planet sample is significantly larger with $\var{N_M}\pm \var{N_M_err}$ planets compared to $\var{N_FGK}\pm \var{N_FGK_err}$ planets in the FGK sample.
No significant detection is possible in the pure FGK sample, independent of the assumed geophysical parameters.
In the M~dwarf sample, the evidence threshold is reached around $f_\mathrm{rgh} \sim 0.2$, similar to the case above where all spectral types are considered.
